### 2015-02-27 icfp15-servo/servo.tex
Servo interop section

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.
Monomorphization, as in MLton~\cite{mlton}, instead instantiates each polymorphic code block at each of the types
it is applied agains, and providing predictable output code to developers, at the cost of some code duplication.
This approach, combined with whole-program compilation, has some negative impacts in compilation speed but has resulted
in Rust code that easily matches the sequential speed of its \Cplusplus{} analog without requiring the Servo developers
to become compiler experts.

@@

\subsection{Language interoperability}
Clean C interop, both for calling and being called were key for integration with various C libraries and bootstrapping larger projects that we intend to eventually be written entirely in Rust, but not initially. This support also means being able to define structs with exact layout expected by C for APIs like that, intead of writing shim layers.

Would like to have full C++ inlining.

That said, interoperating with native C code often requires raw access to memory allocated in Rust or returns
pointers to memory that was allocated by the C libraries.
Translations to and from these raw pointers requires the use of unsafe code.

>>>>

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.
Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied agains, and providing predictable output code to developers, at the cost of some code duplication.
This approach, combined with whole-program compilation~\cite{weeks:whole-program-mlton}, has a negative impact on
compilation speed but has resulted in Rust code that easily matches the sequential speed of its \Cplusplus{} analog
without requiring the Servo developers to become compiler experts.

@@

\subsection{Language interoperability}
Rust has nearly complete interoperability with C code, both exposing code to and using code from
C programs.
This support has allowed us to smoothly integrate with many browser libraries, which has been
critical for bootstrapping a browser without rewriting all of lower-level libraries immediately, such as
graphics rendering code, the JavaScript engine, font selection code, etc.
\tblref{servo-loc} shows the breakdown between current lines of Rust code (including generated code that
handles interfacing with the JavaScript engine) and C code.
This table also includes test code, though the majority of that code is in HTML and JavaScript.
\begin{table}
  \begin{center}
    \begin{tabular}{r || r}
      Language & Lines of Code \\
      \hline
      C or \Cplusplus{} & 1,678,427 \\
      Rust & 410,817 \\
      HTML or JavaScript & 217,827 \\
    \end{tabular}%
  \end{center}%
  \caption{Lines of code in Servo}
  \label{servo-loc}
\end{table}

There are two limitations in the language interoperability that pose challenges for Servo today.
First, Rust cannot currently expose varargs-style functions to C code.
Second, Rust cannot compile against \Cplusplus{} code.
In both cases, Servo uses C wrapper code to call into the code that Rust cannot directly
reach.
While this approach is not a large problem for varargs-style functions, it defeats many of the
places where the \Cplusplus{} code has been crafted to ensure the code is inlined into the caller,
which results in degraded performance.
